Continuous Deductive Reasoning over Streamed RDF Data

### 2015-11-10 article.tex
clarified definition of satisfaction in TA Datalog and made wording consistent regarding the open-ended semantics of truth in fact-instances.

When considering a TA-Datalog program, in which there may be many
distinct instances of a given fact, temporal information regarding
when instances of that fact became true would be lost if such a
program were to entail sets of plain facts without temporal
annotations. We therefore extend the model theory for Datalog from
\refdef{continuous datalog: datalog program} by considering the
Herbrand base of all fact instances $\hat{H}^T$ (all facts with all
instants). As before, a fact instance $\langle e, t_e \rangle$ is 
entailed by a program $\Pi^T$ iff each interpretation
$\mathcal{I}^T \subseteq \hat{H}^T$ satisfying every
clause (rule in $R$ and fact instance in $F^T$) in $\Pi^T$ also satisfies
$\langle e, t_e \rangle$, written $\Pi^T \models \langle e, t_e \rangle$.
As with facts in positive Datalog, we say that an interpretation
$\mathcal{I}^T$ satisfies a fact instance $\langle f, t \rangle \in F^T$
iff $\langle f, t \rangle \in \mathcal{I}^T$.  We say the interpretation
$\mathcal{I}^T$ satisfies a rule
$(r = \alpha_0 \Leftarrow \alpha_1 \land \ldots \land \alpha_n) \in R$ iff,
given $\{\langle a_1, t_1 \rangle, \dots, \langle a_n, t_n \rangle\} \subseteq \mathcal{I}^T$
such that $\{a_1, \dots, a_n\}$ satisfies the body of $r$,
$\langle a_0, t_0 \rangle \in \mathcal{I}^T$ (where $t_0 = max(t_1,\dots,t_n)$,
see \refdef{continuous datalog: instance justification}) such that
$\{a_0, \dots, a_n\}$ satisfies $r$.  The {\em entailment of the TA-Datalog
program} $\Pi^T$ is therefore:
\[ cons^T(\Pi^T) = \{ \langle e, t_e \rangle \in \hat{H}^T \mid
\Pi^T \models \langle e, t_e \rangle \} \]

@@

For an entailed fact instance $\langle e, t_e\rangle$ and a
justification of that fact instance $J^T$, the instant $t_e$ at which
$\langle e, t_e \rangle$ becomes true is, intuitively, taken to be the
latest instant at which any of the fact instances in the justification
became true:
\[ t_e = max(\{ t \mid \langle j, t \rangle \in J^T \}) \]

>>>>

When considering a TA-Datalog program, in which there may be many
distinct instances of a given fact, temporal information regarding
when instances of that fact became true would be lost if such a
program were to entail sets of plain facts without temporal
annotations. We therefore extend the model theory for Datalog from
\refdef{continuous datalog: datalog program} by considering the
Herbrand base of all fact instances $\hat{H}^T$ (all facts with all
instants). As before, a fact instance $\langle e, t_e \rangle$ is 
entailed by a program $\Pi^T$ iff each interpretation
$\mathcal{I}^T \subseteq \hat{H}^T$ satisfying every clause in $\Pi^T$
also satisfies $\langle e, t_e \rangle$, written
$\Pi^T \models \langle e, t_e \rangle$.  However, we must extend the
definition of satisfaction from that for positive Datalog to incorporate
the notion of time.  The extension must support the satisfaction of the
rules $R$ of a TA-Datalog program by sets of fact instances, despite the
former containing no notion of fact annotation, and must additionally
uphold the semantics of temporal annotation described in
\refdef{continuous datalog: fact instance}.

To this end, we say that an interpretation $\mathcal{I}^T$ satisfies a
fact instance $\langle f, t \rangle$ iff
$\langle f, t \rangle \in \mathcal{I}^T$, as in positive Datalog.
We also say that the interpretation $\mathcal{I}^T$ satisfies a rule
$r = \alpha_0 \Leftarrow \alpha_1 \land \ldots \land \alpha_n$ iff, for
every \emph{extended variable substitution} $\bar{s}$ by which
$\{\langle \bar{s}\left( \alpha_1 \right), t_1 \rangle, \dots, \langle \bar{s}\left( \alpha_n \right), t_n \rangle\} \subseteq \mathcal{I}^T$, there exists a fact instance
$\langle \bar{s}\left( \alpha_0 \right), max\left(t_1, \dots, t_n \right) \rangle \in \mathcal{I}^T$.
The meaning of the restriction on the instant associated with the instance of
$\bar{s}\left( \alpha_0 \right)$ in $\mathcal{I}^T$ is better described in terms
of an entailment and its time-annotated justification, as per
\refdef{continuous datalog: instance justification}.

The {\em entailment of the TA-Datalog program} $\Pi^T$ is therefore:
\[ cons^T(\Pi^T) = \{ \langle e, t_e \rangle \in \hat{H}^T \mid
\Pi^T \models \langle e, t_e \rangle \} \]

@@

For an entailed fact instance $\langle e, t_e\rangle$ and a
justification of that fact instance $J^T$, the instant $t_e$ from which
$\langle e, t_e \rangle$ is recognised to be true is, intuitively, taken to be the
latest instant from which any of the fact instances in the justification
became true:
\[ t_e = max(\{ t \mid \langle j, t \rangle \in J^T \}) \]
